---
title: "6-2. One-Sample Wilcoxon Signed Rank Test (non-parametric)"
output: html_document
---

### 1. What’s one-sample Wilcoxon signed rank test?

  - The one-sample Wilcoxon signed rank test is a non-parametric alternative to one-sample t-test when the data cannot be assumed to be normally distributed.
        
  - It’s used to determine whether the median of the sample is equal to a known standard value (i.e. theoretical value).

  - *Note* that, 
    + the data should be distributed symmetrically around the median.
    + In other words, there should be roughly the same number of values above and below the median.

***

### 2. Research questions and statistical hypotheses

  - Typical research questions are:

    + whether the median (m) of the sample is **equal** to the theoretical value (m~0~)?
    + whether the median (m) of the sample is **less than** to the theoretical value (m~0~)?
    + whether the median (m) of the sample is **greater than** to the theoretical value(m~0~)?
<p/>
  - In statistics, we can define the corresponding null hypothesis(H~0~) as follow:
        
    + H~0~ : m = m~0~
    + H~0~ : m ≤ m~0~
    + H~0~ : m ≥ m~0~
<p/>
  - The corresponding alternative hypotheses(H~1~) are as follow:

    + H~1~ : m ≠ m~0~ (different)
    + H~1~ : m > m~0~ (greater)
    + H~1~ : m < m~0~ (less)
<p/>
  - *Note* that:
  
    + H~0~ are called `two-tailed` tests if equal or different
    + H~1~ are called `one-tailed` tests if greater or less

***

### 3. Visualize your data and compute one-sample Wilcoxon test in R

  - Install `{ggpubr}` package for data visualization:
    + You can draw R base graphs as described at this link: R base graphs.
    + Here, we’ll use the `{ggpubr}` package for an easy ggplot2-based data visualization

```{r code1, eval=FALSE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# install.packages("ggpubr")
```

  - R function to compute one-sample Wilcoxon test
    + To perform one-sample Wilcoxon-test, the R function `wilcox.test{stats}` can be used as follow:

```{r code2, eval=FALSE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
wilcox.test(x, mu = 0, alternative = "two.sided")
```
  -
    + `x` : a numeric vector containing your data values
    + `mu` : the theoretical mean/median value. Default is 0 but you can change it.
    + `alternative` : the alternative hypothesis. 
      Allowed value is one of `two.sided` (default), `greater` or `less`.
<p/>
  - Import your data into R
    + Here, we’ll use an example data set containing the weight of 10 mice.

>We want to know, if the median weight of the mice differs from 25g?

```{r code3, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
set.seed(1234)

( 
  my_data <- data.frame( 
                name = paste0(rep("M_", 10), 1:10),
                weight = round(rnorm(10, 20, 2), 1)
              )
)
```

  - Check your data

```{r code4, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# Print the random sampled 10 rows of the data
library(dplyr)
sample_n(tbl = my_data, size = 10, replace = FALSE)

# Statistical summaries of weight
summary(my_data$weight)
```
  -
    + `Min.` : the minimum value
    + `1st Qu.` : The first quartile. 25% of values are lower than this.
    + `Median`  : the median value. Half the values are lower; half are higher.
    + `3rd Qu.` : the third quartile. 75% of values are higher than this.
    + `Max.` : the maximum value
<p/>
  - Visualize your data using box plots

```{r code5, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE, fig.width=3, fig.height=3.5}
library(ggpubr)

ggboxplot(my_data$weight, ylab = "Weight (g)", xlab = FALSE, ggtheme = theme_minimal(), fill='lightgray')
```

  - Compute one-sample Wilcoxon test

>We want to know, if the average weight of the mice differs from 25g (two-tailed test)?

```{r code6, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# One-sample wilcoxon test
( res <- wilcox.test(my_data$weight, mu = 25) )

# print only the p-value
res$p.value
```

  - The p-value of the test is 0.005793, which is less than the significance level alpha = 0.05.  
    We can reject the null hypothesis and conclude that the average weight of the mice is  
    significantly different from 25g with a p-value = 0.005793.
<p/>
  - Note that:  
  
    + if you want to test whether the median weight of mice is less than 25g (one-tailed test), 
    type this :

```{r code7, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
wilcox.test(my_data$weight, mu = 25, alternative = "less")
```
  -
    + Or, if you want to test whether the median weight of mice is greater than 25g (one-tailed test), type this :
        
```{r code8, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
wilcox.test(my_data$weight, mu = 25, alternative = "greater")
```

